---
title: "iteration and list columns"
author: "Kennedy Wade"
date: "2025-10-28"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(tibble)
```

## lists 
In R, vectors are limited to a single data class – all elements are characters, or all numeric, or all logical. Trying to join the following vectors will result in coersion, as would creating vectors of mixed types.
```{r}
vec_numeric = 5:8
vec_char = c("My", "name", "is", "Kennedy")
vec_logical = c(TRUE, TRUE, TRUE, FALSE)
```


 Lists provide a way to store anything you want. This flexibility is great, but is offset by a certain … clunkiness. Lists contain indexed elements, and the indexed elements themselves be scalars, vectors, or other things entirely.
 
```{r}
l = list(
  vec_numeric = 5:8,
  mat         = matrix(1:8, 2, 4),
  vec_logical = c(TRUE, FALSE),
  summary     = summary(rnorm(1000)))
l
```


##  Makiing a list of these different dataframes
```{r}
list_norms = 
  list(
    a = rnorm(20, 3, 1),
    b = rnorm(20, 0, 5),
    c = rnorm(20, 10, .2),
    d = rnorm(20, -3, 1)
  )
```


copy and paste the function from last time 
```{r}
mean_and_sd = function(x){
  
  
   if (!is.numeric(x)) {
     stop( "The input x should be numeric")
   }
   
  if (length(x) <5) {
    stop("Only computw z scores when the input has 5 or more numbers")
  }
  
 mean_x = mean(x, na.rm = TRUE)
 sd_x = sd(x, na.rm = TRUE)
 
 tibble(
   mean = mean_x,
   sd = sd_x
 )
 
}
x_vec = rnorm(25, mean = 5, sd = 3)
mean_and_sd(x_vec)
```


We can apply the mean_and_sd function to each element of list_norms using the lines below.
```{r}
mean_and_sd(list_norms[[1]])
mean_and_sd(list_norms[[2]])
mean_and_sd(list_norms[[3]])
mean_and_sd(list_norms[[4]])

```


define an output list with the same number of entries as my target dataframe; a sequence to iterate over; and a for loop body that applies the mean_and_sd function for each sequence element and saves the result.
```{r}
output = vector("list", length = 4)

for (i in 1:4) {
  output[[i]] = mean_and_sd(list_norms[[i]])
}
```


use `map` to do the same thing: The map functions in purrr try to make the purpose of your code clear. Compare the loop above to the line below.
```{r}
output = map(list_norms, mean_and_sd)

output = map(list_norms, median) 
```


check out some `map` variants 
```{r}
map_dfr(list_norms, mean_and_sd, .id = "sample")

map_dfr(list_norms, median)
```


try to out a lsit into a datframe
```{r}
listcol_df = 
  tibble(
   name = c("a","b", "c", "d"),
   sample = list_norms 
  )
  
```

did this work????
```{r}
pull(listcol_df, name)
pull(listcol_df, "sample")
```

yes it did!!!!!

can i apply `mean_and_sd`
```{r}
mean_and_sd(pull(listcol_df, sample)[[1]])
mean_and_sd(pull(listcol_df, sample)[[2]])
mean_and_sd(pull(listcol_df, sample)[[3]])
mean_and_sd(pull(listcol_df, sample)[[4]])
```


iterate using `map`
```{r}
map(pull(listcol_df, sample),mean_and_sd)
```


addiing a column...

```{r}
listcol_df |>
  mutate(
    summary = map(sample, mean_and_sd)
  )
```

```{r}

listcol_df <- listcol_df |>
  mutate(summary = map(sample, mean_and_sd))

listcol_df |> 
  select( -sample) |> 
  unnest(summary)
```

## Revisiting NSDUH\

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

data_marj = 
  nsduh_html |> 
  html_table() |> 
  nth(1) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```


```{r}
nsduh_table <- function(html, table_num) {
  
  table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent)) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  table
}

nsduh_table(nsduh_html, table_num = 1)
nsduh_table(nsduh_html, table_num = 2)
nsduh_table(nsduh_html, table_num = 3)

```


try this with a `for` loop 
```{r}
output = vector("list", length = 3)

for (i in 1:3) {
  output[[i]] = nsduh_table(nsduh_html, i)
}
```


do this with `map`
```{r}
map(1:3, nsduh_table, html = nsduh_html)
```


do all this in a dtaframe 
```{r}
nsduh_df = 
  tibble(
    name = c("marj year", "marj month", "marj year"),
    number = 1:3
  ) |>
  
  mutate(
    table = map(number, nsduh_table,html = nsduh_html)
  ) |>

  unnest(table)
  
```

## lets look at the weather data 

```{r}
library(p8105.datasets)
data("weather_df")
```


```{r}
weather_df |>
  filter(name =="CentralPark_NY") |>
  ggplot(aes( x = tmin, y = tmax)) +
  geom_point() + 
  geom_smooth(method = "lm")
```


lets do regression
```{r}
weather_df |>
  filter(name =="CentralPark_NY") |>
 lm(tmax ~ tmin, data = _) 

 weather_df |>
  filter(name =="Molokai_HI") |>
 lm(tmax ~ tmin, data = _) 

 weather_df |>
  filter(name =="Waterhole_WA") |>
 lm(tmax ~ tmin, data = _) 
```


lets iterate differently...
```{r}
weather_nest <-weather_df |> 
  nest(data = date:tmin)
```


```{r}
lm(tmax ~tmin, data = pull(weather_nest, data)[[1]])
lm(tmax ~tmin, data = pull(weather_nest, data)[[2]])
lm(tmax ~tmin, data = pull(weather_nest, data)[[3]])

```

do this using `map`
```{r}
weather_lm =function(df) {
  
  lm(tmax ~ tmin, data = df)

}
```

```{r}
map(pull(weather_nest, data), weather_lm)
```


```{r}
weather_nest |>
  mutate(
     lm_fits = map(data, weather_lm)
  )
```


