---
title: "writing_functions"
author: "Kennedy Wade"
date: "2025-10-23"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)

set.seed(1)
```

If I want to repeat this (admittedly simple) process for lots of samples, I might want to have a function that takes the sample as an argument, computes the vector of Z scores in the body, and returns the result. I define such a function below.


write a function t compute z scores... 
```{r}
z_scores = function(x) {
  
   if (!is.numeric(x)) {
     stop( "The input x should be numeric")
   }
   
  if (length(x) <5) {
    stop("Only compute z scores when the input has 5 or more numbers")
  }
    
  z = (x - mean(x)) / sd(x)
  return(z)
}
```

## My first function

The best way to build up a function is to start with code you’ve written outside a function. To see how this might work, I’ll start with a simple example: the chunk below takes a sample from a normal distribution and then computes the vector of Z scores for the sample.

```{r}
x_vec = rnorm(25, mean = 5, sd = 3)

(x_vec - mean(x_vec)) / sd(x_vec)

num_vec = rnorm(123, mean = 14, sd = 0.4)

z_scores(x = num_vec)
```


lets break our function...

```{r, error = TRUE}

z_scores(3)
z_scores("my name is Kennedy")
```


## Lets compute some stuff 

lets compute and return the mean and sd of a numeric vector 

```{r}
mean_and_sd = function(x){
  
  
   if (!is.numeric(x)) {
     stop( "The input x should be numeric")
   }
   
  if (length(x) <5) {
    stop("Only computw z scores when the input has 5 or more numbers")
  }
  
 mean_x = mean(x, na.rm = TRUE)
 sd_x = sd(x, na.rm = TRUE)
 
 tibble(
   mean = mean_x,
   sd = sd_x
 )
 
}

mean_and_sd(x_vec)

```


## Make up data...

lets "simulate" some data

```{r}
sim_data = tibble(
  x = rnorm(30, mean = 3, sd = 2)
)

sim_data |> 
  summarize(
    mu_hat = mean(x),
    sigma_hat = sd(x)
  )
```


write a function to do simulations

the inputs are:

*`n_subj` is number of subjects 
*`mu` is the true data 
* `sigma` is the true sd

```{r}
sim_mean_sd = function(n = 30, mu = 3, sigma = 2) {
  sim_data = tibble(
    x = rnorm(n = n, mean = mu, sd = sigma)
  )
  
  sim_data |> 
    summarize(
      mu_hat = mean(x),
      sigma_hat = sd(x)
    )
}

sim_mean_sd()
sim_mean_sd(n = 50, mu = 10, sigma = 5)

```

lets run this function

```{r}
sim_mean_sd(mu = 48, sigma = 50)

```

Import the data and turn this into a function

```{r}
lotr_import = function(cell_range, movie_title) {
  df = readxl::read_excel("data/LotR_Words.xlsx", range = cell_range) |>
    mutate(movie = movie_title)
  
  return(df)
}

# Now use the function
fellowship = lotr_import(cell_range = "B3:D6", movie_title = "Fellowship")
two_towers = lotr_import(cell_range = "F3:H6", movie_title = "Two Towers")
return_of_the_king = lotr_import(cell_range = "J3:L6", movie_title = "Return of the King")

# Combine them into one tibble
lotr_df = bind_rows(fellowship, two_towers, return_of_the_king) |>
  janitor::clean_names() |>
  pivot_longer(
    female:male,
    names_to = "sex",
    values_to = "words"
  ) |>
  mutate(race = str_to_lower(race)) |>
  select(movie, everything())

fellowship_ring = readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") |>
  mutate(movie = "fellowship_ring")

two_towers = readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") |>
  mutate(movie = "two_towers")

return_of_the_king = readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") |>
  mutate(movie = "return_of_the_king")

lotr_df = bind_rows(fellowship_ring, two_towers, return_of_the_king) |>
  janitor::clean_names() |>
  pivot_longer(
    female:male,
    names_to = "sex",
    values_to = "words") |> 
  mutate(race = str_to_lower(race)) |> 
  select(movie, everything())

```



```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

data_marj = 
  nsduh_html |> 
  html_table() |> 
  nth(1) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```

write an import function

```{r}
nsduh_table <- function(html, table_num, table_name) {
  
  table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent),
      name = table_name) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  table
  
}

#use this to get a few different tables and combine the results.

nsduh_results = 
  bind_rows(
    nsduh_table(nsduh_html, 1, "marj_one_year"),
    nsduh_table(nsduh_html, 4, "cocaine_one_year"),
    nsduh_table(nsduh_html, 5, "heroin_one_year")
  )

```



